Asymmetric relationships and their effects on coexistence

### 2023-09-26 paper/RefereeResponses2.tex

\noindent \emph{Dear Editors,} \\

\noindent \emph{Thank you for the decision of ``minor revisions''. We have 
fully addressed the referee and editor comments below and in the manuscript.
We hope the paper can now be rapidly formally accepted. } \\

\noindent \emph{We have replied to referee comments below in italic, with each response preceded
by ``***'', to make our responses easier to find.} \\

\noindent \emph{We submit the following pdf files: The main text pdf, including all figures, embedded 
next to their captions for readability; 
the Supporting Information; a version of the main text with changes tracked; and all figures as 
separate pdfs, for the production process. The document showing 
the changes we made does not include links such as citations and
references to figures, but it does accurately represent the changes made to the text.
We also submit several latex and supporting files, for the production process.} \\

### 2023-09-28 paper/Paper.Rnw
deleted a superfluous comment and ammended the responses
adjustments required by the publisher to conform with data availability requirements

\noindent \textbf{Running head:} Asymmetric relationships and coexistence \\
\noindent \textbf{Key words:} asymmetric tail association, biodiversity, coexistence, competition, copula, modern coexistence theory, niche differentiation, storage effects \\
\noindent \textbf{Type of article:} Letter\\
\noindent \textbf{Words in Abstract:} 150; \textbf{Words in main text:} 4998; \textbf{Words in text boxes:} 0\\
\noindent \textbf{Number of references:} 39; \textbf{Number of figures:} 5; \textbf{Number of tables:} 1; \textbf{Number of text boxes:} 0\\
\noindent \textbf{Corresponding author:}
Daniel C. Reuman,
Department of Ecology and Evolutionary Biology,
University of Kansas,
2101 Constant Ave,
Lawrence, KS 66047;
Phone: 626-560-7084 \\
\noindent \textbf{Author emails:} pimsupaa@uoregon.edu, reuman@ku.edu\\
\noindent \textbf{Statement of Authorship:} The authors jointly carried out all parts of the study.\\
\noindent \textbf{Data accessibility statement:} All codes associated with this study have been released in a 
public repository.

@@

All computations were done in R on a laptop running Ubuntu Linux. 
Complete computer codes for the project are stored at 
\url{https://github.com/jasminalbert/AsymmetricRelationshipsAndCoexistence.git}.

@@

@@

The authors thank Shyamolina Ghosh for suggesting the research problem, and Lawrence Sheppard, Max Castorani,
Jon Walter, Jude Kastens, Vadim Karatayev, Emily Arsenault, Jordan Rodriguez, Lauren Hallett, 
Madelon Case, Chhaya Werner, Lina Aoyama, Steven Haring, Carmen Watkins,
Robin Snyder, and three anonymous referees and the editor for helpful feedback. 
The authors especially thank Steve Ellner for key advice at several watershed moments in the development of the project. 
The authors were partly supported by US National
Science Foundation grants 1714195 and 2023474, the James S McDonnell Foundation, the California Department of Fish and
Wildlife Delta Science Program, and the Alexander von Humboldt Foundation.

@@

\begin{figure}
\includegraphics[width=\textwidth]{../results_figs/fig1_a-h.pdf}
\caption{Examples motivating our study of the effects of asymmetric tail associations (ATAs) on competition
and coexistence. (a-c) Bivariate random variables with standard normal marginals showing
alternative patterns of association in the left and right parts of the distributions. 
The two variables in (a) are more strongly related to each
other in the left parts of the distributions, termed ``left-tail association;'' those in (b) 
are symmetrically associated; and those in (c) 
are more strongly related to each other in the right parts, termed ``right-tail association''. All three cases have 
the same Pearson correlation, $\rho$, up to sampling variation, so Pearson correlation does not identify these 
very different patterns of association. For (a, c), we used asymmetrically tail associated cases which are perfectly 
correlated in the lower or upper halves of the distributions, though it is also possible to generate asymmetrically 
tail associated noise with imperfect correlations in both halves of the distributions [\citealp{Nelsen2006,Ghosh_2020_AER};
see also (d,e)]. 
%The variables $(B_1,B_2)$ described in the Introduction and Methods were obtained by generating noise as 
%pictured and then transforming the marginals as specified in Methods.
(d,e) Two examples of ATAs found in nature. Axes are population densities of two plankton species, 
\emph{Ceratium fusus} and \emph{Ceratium furca}, sampled in successive years in two of the 15 locations 
in the North Sea considered by \citep{Ghosh_2020_ecolevol}. ATAs were statistically
significant and differed by location, with associations being primarily in the left tails in 
panel d, and primarily in the right tails in panel e (SI section \ref{SI-SIsect:planktonstats}). 
One can see these asymmetries by noting that the open points in panel d, which are in the left halves of the
distributions, are more tightly clustered around the line $y=x$ than are the closed points,
which are in the right halves of the distributions; and \emph{vice versa} for panel e.
(f-h) Lottery model population simulations % using $(B_1,B_2)$ 
obtained by transforming the respectively corresponding variables (a-c; see Introduction).
\label{fig:pedagog1}}
\end{figure}

@@

%***DAN: Jasmin, this currently pulls in a pdf you sent me, stored temporarily on my drive in the indicated 
%location. Once the code produces fig 5 on a normal run, we'll have to change the below to suck that
%version of the figure up instead of this one.
\begin{figure}
\includegraphics[width=\textwidth]{../results_figs/fig5_new.pdf}
\caption{See next page for caption.}  \label{fig:fig5}
\end{figure}

>>>>

\noindent \textbf{Running head:} Asymmetric relationships and coexistence \\
\noindent \textbf{Key words:} asymmetric tail association, biodiversity, coexistence, competition, copula, modern coexistence theory, niche differentiation, storage effects \\
\noindent \textbf{Type of article:} Letter\\
\noindent \textbf{Words in Abstract:} 150; \textbf{Words in main text:} 4998; \textbf{Words in text boxes:} 0\\
\noindent \textbf{Number of references:} 40; \textbf{Number of figures:} 5; \textbf{Number of tables:} 1; \textbf{Number of text boxes:} 0\\
\noindent \textbf{Corresponding author:}
Daniel C. Reuman,
Department of Ecology and Evolutionary Biology,
University of Kansas,
2101 Constant Ave,
Lawrence, KS 66047;
Phone: 626-560-7084 \\
\noindent \textbf{Author emails:} pimsupaa@uoregon.edu, reuman@ku.edu\\
\noindent \textbf{Statement of Authorship:} The authors jointly carried out all parts of the study.\\
\noindent \textbf{Data accessibility statement:} All codes associated with this study have been released in a 
public repository.

@@

All computations were done in R on a laptop running Ubuntu Linux. 
The only data used were the phytoplankton time series of Fig. 1, taken from
\url{https://doi.org/10.5061/dryad.rq3jc84}.
Complete computer codes and data are stored publicly as indicated in the Data Availability Statement.

@@

\section{Author contributions}

Both authors contributed about equally to all parts of the paper.

@@

The authors thank Shyamolina Ghosh for suggesting the research problem, and Lawrence Sheppard, Max Castorani,
Jon Walter, Jude Kastens, Vadim Karatayev, Emily Arsenault, Jordan Rodriguez, Lauren Hallett, 
Madelon Case, Chhaya Werner, Lina Aoyama, Steven Haring, Carmen Watkins,
Robin Snyder, and three anonymous referees and the editor for helpful feedback. 
The authors especially thank Steve Ellner for key advice at several watershed moments in the development of the project.

\section{Funding information}

The authors were partly supported by US National
Science Foundation grants 1714195 and 2023474, the James S McDonnell Foundation, the California Department of Fish and
Wildlife Delta Science Program, and the Alexander von Humboldt Foundation.

\section{Data availability statement}

All code and data for the project are available at \url{https://zenodo.org/record/XXX}
%DAN: update when you know
and \url{https://github.com/jasminalbert/AsymmetricRelationshipsAndCoexistence}.

@@

\begin{figure}
\includegraphics[width=\textwidth]{../results_figs/fig1_a-h.pdf}
\caption{Examples motivating our study of the effects of asymmetric tail associations (ATAs) on competition
and coexistence. (a-c) Bivariate random variables with standard normal marginals showing
alternative patterns of association in the left and right parts of the distributions. 
The two variables in (a) are more strongly related to each
other in the left parts of the distributions, termed ``left-tail association;'' those in (b) 
are symmetrically associated; and those in (c) 
are more strongly related to each other in the right parts, termed ``right-tail association''. All three cases have 
the same Pearson correlation, $\rho$, up to sampling variation, so Pearson correlation does not identify these 
very different patterns of association. For (a, c), we used asymmetrically tail associated cases which are perfectly 
correlated in the lower or upper halves of the distributions, though it is also possible to generate asymmetrically 
tail associated noise with imperfect correlations in both halves of the distributions [\citealp{Nelsen2006,Ghosh_2020_AER};
see also (d,e)]. 
%The variables $(B_1,B_2)$ described in the Introduction and Methods were obtained by generating noise as 
%pictured and then transforming the marginals as specified in Methods.
(d,e) Two examples of ATAs found in nature. Axes are population densities of two plankton species, 
\emph{Ceratium fusus} and \emph{Ceratium furca}, sampled in successive years in two of the 15 locations 
in the North Sea considered by \cite{Ghosh_2020_ecolevol} and \cite{Sheppard2019}. ATAs were statistically
significant and differed by location, with associations being primarily in the left tails in 
panel d, and primarily in the right tails in panel e (SI section \ref{SI-SIsect:planktonstats}). 
One can see these asymmetries by noting that the open points in panel d, which are in the left halves of the
distributions, are more tightly clustered around the line $y=x$ than are the closed points,
which are in the right halves of the distributions; and \emph{vice versa} for panel e.
(f-h) Lottery model population simulations % using $(B_1,B_2)$ 
obtained by transforming the respectively corresponding variables (a-c; see Introduction).
\label{fig:pedagog1}}
\end{figure}

@@

\begin{figure}
\includegraphics[width=\textwidth]{../results_figs/fig5_new.pdf}
\caption{See next page for caption.}  \label{fig:fig5}
\end{figure}
